Studying public transit API query logs\\ to get an indication of travel flows

### 2016-01-06 paper.tex
Shifted around chapters 6 and 7, extended conclusion with future work.

The iRail API is a XML/JSON HTTP API where each request on \url{http://api.irail.be} returns a response that can be used directly in an end-user app. It contains four features: 
\begin{enumerate}
\item \emph{Route Planning} (\url{http://api.irail.be/connections/{?from,to,date,time}}) plans a route from one station to another, taking into account a preferred start time.
\item \emph{Next Departures} (\url{http://api.irail.be/liveboard/{?station,date,time}}) shows the next train departures in a certain station, useful for creating quick overviews.
\item \emph{Trip Status} (\url{http://api.irail.be/vehicle/{?id,date}}) returns the stations this train passes by on the day of the request.
\item \emph{List of all stations} (\url{http://api.irail.be/stations/}) gives an overview of all stations Belgian trains can arrive at.
\end{enumerate}

@@

@@

%%% EXPLAIN:
%%%
The expressiveness of a server affects the way logs can be gathered.
We have shown in previous sections that the query logs of a \emph{high expressive server}, such as the iRail API, are interesting: each request contains an entire query which can be interpreted as a travel intention.
The possibility that two requests are exactly the same is low, as there are many URLs which can be requested.
Nevertheless, we cannot fully guarantee that each HTTP GET request to the server will trigger a log entry, due to caching mechanisms on the Web~\cite{fielding}, which might lead to a false representation.

@@

In \cref{fig:LDFAxis2} we illustrate these two options as two extremes, with other options that are yet to be discovered.
When a \emph{rather low expressive} server only allows to set e.g., a departure station and a departure time, then the server cannot log the arrival station, yet the client is still able to plan a route by executing the algorithm on the client-side~\cite{lc}.
We would be able to fully rely on the query logs if the expressivity would be maximal (extreme right) and caching would be turned off.
Turning off caching may work in a private setting where you know how much and what kind of queries you can expect, yet on the Web, it will not scale up.\todo{PC: Just being devil's advocate here: is caching really necessary to scale sufficiently?}

\section{Publishing query logs}
\label{sec:publishingquerylogs}

In order to publish this data, we chose a newline-based format: each new line contains a new query object encoded in JSON. 
We call this a \emph{JSON-stream}, as the JSON objects can be read one per one.
Unix commands such as \emph{grep} work well with this format.
We chose this over CSV as we have more flexibility towards changing the object properties in the future.
Next, we chose to add an object to the stream which only contains an object with key \emph{@context}.
This is the JSON-LD context of all the objects in the, JSON-stream. Subsequentially, we call this a \emph{JSONLD-stream} and this has been formalized at \emph{https://github.com/pietercolpaert/jsonld-stream}.

%As a new version of the logs are put at \url{http://api.irail.be/logs/} only each second, we have added cache-headers to the document with max-age equal to 1 second.
%\todo{Also in USEWOD2014 there was a paper on LDF and how we could know which SPARQL queries needed to be solved}

\section{Future Work}

Query logs are thus an indication of travel intentions. 
Yet we can identify some hiatuses in query logs as travel intentions:
\begin{enumerate}
  \item Caching may mask peak hours, as \ldots
  \item Multiple requests may be needed to cover a travel intention
  \item \ldots
\end{enumerate}

In order to create a better representation of travel intentions, we would recommend apps to do a POST-request instead of a

\section{Conclusion}
\label{sec:conclusion}

The contribution of this paper is twofold.
For the first time, we studied query logs to find travel patterns.
Due to lack of information we could not evaluate our results, yet the visualizations look promising...
There are obvious caveats associated with the use of such data as proxy for actual statistical counts. 
For instance, the data only captures an intention of a passenger: it's unsure whether the person actually took the train.
Furthermore, different requests could be done by one person with an intention to travel.
On top of that, our sample used for this paper can only represent a maximum of 0.4\% of the daily traffic flow.

The second contribution is our position that we suggest to decouple the query logs from the query execution.

In future work, iRail will try to use these to predict train congestion.
% Fix spacing after References header (as line 1308 of sig-alternate.cls breaks it)
\let\oldsection\section
\renewcommand{\section}[2][1]{\oldsection{#1}\vspace{-3pt}}

>>>>

\subsection{Historical logs}

The iRail API is an XML/JSON HTTP API: each request on \url{http://api.irail.be} returns a response that can be used directly in an end-user app.
It contains four features: 
\begin{enumerate}
\item \emph{Route Planning} (\url{http://api.irail.be/connections/{?from,to,date,time}}) plans a route from one station to another, taking into account a preferred start time.
\item \emph{Next Departures} (\url{http://api.irail.be/liveboard/{?station,date,time}}) shows the next train departures in a certain station, useful for creating quick overviews.
\item \emph{Trip Status} (\url{http://api.irail.be/vehicle/{?id,date}}) returns the stations this train passes by on the day of the request.
\item \emph{List of all stations} (\url{http://api.irail.be/stations/}) gives an overview of all stations Belgian trains can arrive at.
\end{enumerate}

@@

\subsection{Real-time logs}
\label{sec:realtimelogs}

In order to publish this data, we chose a newline-based format: each new line contains a new query object encoded in JSON. 
The JSON objects can be read one per one, so unix commands such as \emph{grep} work well with this format.
We chose this over CSV as we have more flexibility towards changing the object properties in the future.
Next, we chose to add an object to the stream which only contains an object with key \emph{@context}.
This is the JSON-LD context of all the objects in the, JSON-stream. Subsequentially, we call this a \emph{JSONLD-stream} and this has been formalized at \emph{https://github.com/pietercolpaert/jsonld-stream}.

Using this data, everyone can build their own historic query log dataset.
The server costs for iRail to host this dataset stay limited.
Furthermore, this way of publishing data enables research for real-time predictions of travel intentions.
For this paper, we will focus on using the historic logs.
Later, the research can be reproduced with more data.

@@

%%% EXPLAIN:
%%%
The expressiveness of a server affects the way logs can be gathered~\cite{usewod2015}.
We have shown in previous sections that the query logs of a \emph{high expressive server}, such as the iRail API, are interesting: each request contains an entire query which can be interpreted as a travel intention.
The possibility that two requests are exactly the same is low, as there are many URLs which can be requested.
Nevertheless, we cannot fully guarantee that each HTTP GET request to the server will trigger a log entry, due to caching mechanisms on the Web~\cite{fielding}, which might lead to a false representation.

@@

In \cref{fig:LDFAxis2} we illustrate these two options as two extremes, with other options that are yet to be discovered.
When a \emph{rather low expressive} server only allows to set e.g., a departure station and a departure time, then the server cannot log the arrival station, yet the client is still able to plan a route by executing the algorithm on the client-side~\cite{lc}.
We would be able to fully rely on the query logs if the expressivity would be maximal (extreme right) and caching would be turned off.
Turning off caching may work in a private setting where you know how much and what kind of queries you can expect, yet on the open Web, caching helps decrease the CPU load on servers on  peak moments.

\section{Conclusion and future work}
\label{sec:conclusion}

First we studied query logs to find travel patterns in Belgium and can conclude this is up to now, the best representation of travel flows over the railway network in Belgium. %, as no alternatives are available.
We studied data which represent a maximum of 1.66\% of the passengers.
\todo{ALVIN: Visualizing our dataset reveals the complexity of movement on the Belgian rail system and the significance of cities in weekday travel.
Furthermore, we found evidence that Flanders is polycentric, while Walloon traffic converges on Brussels.}

Next, we also concluded that there are obvious caveats associated with the use of such data as proxy for actual statistical counts. 
We identified a couple of gaps:
\begin{enumerate}
  \item The data only captures an intention of a passenger: it's unsure whether the person actually took the train.
  \item Multiple user queries may be needed to cover a travel intention.
  \item Caching may mask peak hours, as many of the same requests in one minute is only stored in the logs once.
\end{enumerate}
Nevertheless, we published the query logs as open data at \url{http://api.irail.be/logs} for other scientists to continue to research possibilities with this dataset.
For instance, predicting trip congestion on the basis of this data, looks promising. %the iRail project is planning to use this data to predict trip congestion on the basis of this data.

Related work~\cite{verborgh2014lonesome} suggested to add additional metadata in the HTTP headers in order to track the user's intention.
While this is an interesting suggestion, it does not overcome the problem of caching.
In order to create a better representation of travel intentions, future work can gather logs by POST-requests with the only purpose to gather analytics of travel intention.
This is a similar approach than what happens on the Web of Documents today with \emph{Piwik} or \emph{Google Analytics}.

% Fix spacing after References header (as line 1308 of sig-alternate.cls breaks it)
\let\oldsection\section
\renewcommand{\section}[2][1]{\oldsection{#1}\vspace{-3pt}}
